Bringing the Power of eBPF to Open vSwitch

### 2018-10-25 paper.tex
Edit 4.4.0 and 4.4.1

\subsection{Performance Evaluation}
All of our performance results use a hardware test bed consisting of
two Intel Xeon E5 2440 v2 1.9GHz servers, each with 1 CPU socket and
8 physical cores with hyperthreading enabled.
One server, the target server, has an Intel 40GbE XL710 single port NIC,
and the other server, the source server, has a Netronome NFP-4000 40GbE device.
The target server runs Linux kernel 4.19-rc4, with i40e driver supporting
the AF\_XDP zero copy mode~\cite{af_xdp_patch_i40e, af_xdp_patch_zc}.
We installed OVS-AFXDP on the target server, and used the source server to
generates 64-byte single UDP flow packets
at the rate of 19~Mpps using the DPDK library.
On the target server, for demonstrating the best performance,
we disabled the Intel Spectre and Meltdown fixes~\cite{jesper_xdp_perf_drop, retpoline}.

For all our experiments, we use a microbenchmark program, called xdpsock,
as the baseline to compare with OVS-AFXDP implementation.
xdpsock is an AF\_XDP sample program doing no packet processing, and can be configured
to do simply dropping all packets, rxdrop, or forwarding the packet to the same port as
it receives, l2fwd.
In our testbed, we measured $19Mpps$ for xdpsock-rxdrop and $17Mpps$ for xdpsock-l2fwd.
For OVS-AFXDP, we conduct similar two experiments, but with OVS's OpenFlow rule installed as
below:
\begin{itemize}
\item OVS-AFXDP rxdrop: Install a single OpenFlow rule to drop every packets,
e.g.,  \texttt{in\_port=eth1, actions=drop}.
\item OVS-AFXDP l2fwd: Install a single OpenFlow rule to forward packet to the same port
as it receives, e.g., \texttt{in\_port=eth1, actions=set\_fields:eth2->in\_port, output:eth1}.
\end{itemize}

Additionally, AF\_XDP provides three operation modes, from slowest to fastest:
skb mode: works on devices using generic XDP~\cite{genericxdp},
driver mode: works on devices with XDP support, and
zero-copy driver mode: works on devices with XDP\_XDP zerocopy support.
For all our experiments, we use the zero-copy driver mode for showing the best performance.

\subsubsection{PMD netdev}
Our initial prototype shows pretty terrible performance, only $0.5Mpps$ for rxdrop.
We observe that when dropping/forwarding packets under OVS-AFXDP, two CPUs show up
100\% utilitization:
\begin{itemize}
\item ovs-vswitchd: This is the process keeps doing the send and receive steps
in Figure~\ref{afxdprx} and~\ref{afxdptx}.
\item ksoftirqd/core\_id: This is the kernel software interrupt thread handling the
incoming packets, triggering XDP program to pass packets to the XSK, and also
processing transmission.
\end{itemize}
\mycomment{Compared to xdpsock, ovs-vswitchd adds on top of it a umempool,
and packets parsing, lookup and action execution.}
We investigate using Linux perf and our first fix is to enable OVS's PMD-mode,
Poll-Mode-Driver, to the afxdp netdev.

In OVS, a non-pmd mode netdev does packet receiption by putting all the receiving
netdev's file descriptors, fds, together and ovs-vswitchd polls them when there is an fd
ready to perform I/O.  As a result, the fd of the XSK is {\em shared} with other fds, and
we observed that the poll system call has pretty high overhead.

Applying OVS's PMD netdev avoids these problems and improves the rxdrop performance
from $0.5Mpss$ to $3Mpps$.  Ideally, receiving packets from XSK should not need to context
switch into kernel as mentioned in steps~\ref{afxdprx}. So an ideal implementation
should show minimal context switches with most of the CPU time stays in userspace. 
When enabling PMD netdev for afxdp, ovs-vswitch can be
configured to create a {\em dedicated} thread for the packet processing.
In OVS-AFXDP, we create a thread for each XSK's receive queue and the thread keeps polling
the RX ring for new packets.  Each round of receive polling processes a batch of
packets, up to a maximum of 32 packets. In addition, we enable 1GB huge memory page support
to reduce the page fault overhead.

@@

Although we allocate one umem per netdev, there might be multiple queues per
netdev sharing the same umem. As a result, the above three data structures
require a mutex lock before accessing umem freelist. Linux perf reports
pthread mutex lock related APIs as ones of the top CPU utilization functions.
We change our design by 1) allocating per-queue umem region and 2)
allocating one pmd thread per queue. As a result, no lock is needed because
each queue has only one thread and its own set of umem elements.

@@

\subsubsection{Batching Send Syscall}
%574.790 M/sec  
We continued measuring the performance of OVS-AFXDP l2fwd and
observed only $4Mpps$, compared to $17Mpps$ baseline xdpsock-l2fwd.
We found that the OVS pmd thread under rxdrop
has much fewer context switches compared to the l2fwd, indicating that
the pmd process spends much more time in kernel space than in userspace.
By using strace, we saw that the OVS-AFXDP l2fwd experiement calls sendto system call
at very high frequeuce, because from Figure~\ref{afxdptx}, we design to check
the success of send (step 4) immediately after issusing send (step 3).
We change the design by calling sendmsg syscall (step 3) only when TX ring is
closed to full, e.g., when 3/4 ring elements have been used.
In specific case, instead of issuing sendmsg syscall for a batch of 32 packets
and making sure they are done, we only issue send when there are 512 outstanding
packets.  With this change, the OVS-AFXDP l2fwd experiment shows around $14Mpps$.

>>>>

\subsection{Performance Evaluation and Optimization}
All of our performance evaluation are conducted on a test bed consisting of
two Intel Xeon E5 2440 v2 1.9GHz servers, each with 1 CPU socket and
8 physical cores with hyperthreading enabled.
The first server, the source server, is served as a traffic generator.
It is equipped with a Netronome NFP-4000 40GbE device, and runs a customized
traffic generator that generates 64-byte single UDP flow packets
at the rate of $19Mpps$ using the DPDK library.
The second server, the target server, is equipped with an Intel 40GbE XL710
single port NIC, and it runs Linux kernel 4.19-rc4 with i40e driver supporting
the AF\_XDP zero copy mode~\cite{af_xdp_patch_i40e, af_xdp_patch_zc}.
We install OVS-AFXDP on the target server, and
we disable the Intel Spectre and Meltdown fixes~\cite{jesper_xdp_perf_drop, retpoline}
for demonstrating the best performance. In addition, we enable 1GB huge
memory page support to reduce the page fault overhead.

For all our experiments, we use a microbenchmark program, called xdpsock,
as the baseline to compare with OVS-AFXDP implementation.
xdpsock is an AF\_XDP sample program that can be configured into two modes.
In the rxdrop mode, xdpsock simply drops all incoming packets without any
packet processing. In the l2fwd mode, xdpsock forwards the packet to the
same port from where it is received.
In our testbed, we measure $19Mpps$ for xdpsock-rxdrop and $17Mpps$ for xdpsock-l2fwd.
For OVS-AFXDP, we conduct two similar experiments, with OVS's OpenFlow rule installed as
below:
\begin{itemize}
\item OVS-AFXDP rxdrop: Install a single OpenFlow rule to drop every packets,
e.g.,  \texttt{in\_port=eth1, actions=drop}.
\item OVS-AFXDP l2fwd: Install a single OpenFlow rule to forward packet to the same port
as it receives, e.g., \texttt{in\_port=eth1, actions=set\_fields:eth2->in\_port, output:eth1}.
\end{itemize}

AF\_XDP can provide the following three operation modes depending on the device
and driver support. We list the three mode from the slowest to the fastest
throughput.
1) skb mode: works on devices using generic XDP~\cite{genericxdp},
2) driver mode: works on devices with XDP support, and
3) zero-copy driver mode: works on devices with XDP\_XDP zerocopy support.
For all our experiments, we use the zero-copy driver mode since it provides the
best performance.

\subsubsection{PMD netdev}
In the beginning, our initial prototype of OVS-AFXDP only provides $0.5Mpps$
throughput for rxdrop.
We observe that when dropping/forwarding packets under OVS-AFXDP, two processes
below can easily reach 100\% CPU utilization:
\begin{itemize}
\item ovs-vswitchd: This is the process that keeps doing the send and receive steps
in Figure~\ref{afxdprx} and~\ref{afxdptx}.
\item ksoftirqd/core\_id: This is the kernel software interrupt thread handling the
incoming packets, triggering XDP program to pass packets to the XSK, and also
processing transmission.
\end{itemize}
\mycomment{Compared to xdpsock, ovs-vswitchd adds on top of it a umempool,
and packets parsing, lookup and action execution.}
With further investigation using Linux perf tools, our first improvement is to
to enable OVS's Poll-Mode-Driver (PMD) mode to the afxdp netdev.

In OVS's non-PMD mode, ovs-vswitchd does packet reception by putting all the
receiving netdev's file descriptors (fd)s together and invokes the poll system
call to determine if any one of the fd is ready to perform I/O.  As a result,
the polling of XSK's fd  is {\em shared} with other fds, and
we observed that the poll system call incurs high overhead.

Applying OVS's PMD netdev avoids these problems and improves the rxdrop performance
from $0.5Mpss$ to $3Mpps$. This performance gain is because of reducing userspace
and kernel space context switch overhead by omitting poll system call.
That is instead of using poll system call, we implement packet reception by using
PMD netdev for AFXDP, so that ovs-vswitchd allocate a dedicated thread for each
XSK's receive queue and the thread keeps polling the RX ring for new packets.
Currently, each round of receive polling processes a batch up to 32 packets.

@@

Although we allocate one umem per netdev, there might be multiple queues per
netdev sharing the same umem. As a result, the above three data structures
require a mutex lock before accessing umem freelist. Linux perf reports
pthread mutex lock related APIs as ones of the top CPU utilization functions.
We change our design by 1) allocating per-queue umem region and 2)
allocating one PMD thread per queue. As a result, no lock is needed because
each queue has only one thread and its own set of umem elements.

@@

\subsubsection{Batching Send Syscall}
%574.790 M/sec  
We continued measuring the performance of OVS-AFXDP l2fwd and
observed only $4Mpps$, compared to $17Mpps$ baseline xdpsock-l2fwd.
We found that the OVS PMD thread under rxdrop
has much fewer context switches compared to the l2fwd, indicating that
the PMD process spends much more time in kernel space than in userspace.
By using strace, we saw that the OVS-AFXDP l2fwd experiement calls sendto system call
at very high frequeuce, because from Figure~\ref{afxdptx}, we design to check
the success of send (step 4) immediately after issusing send (step 3).
We change the design by calling sendmsg syscall (step 3) only when TX ring is
closed to full, e.g., when 3/4 ring elements have been used.
In specific case, instead of issuing sendmsg syscall for a batch of 32 packets
and making sure they are done, we only issue send when there are 512 outstanding
packets.  With this change, the OVS-AFXDP l2fwd experiment shows around $14Mpps$.
